Project Report\\\textsc{DEVS Ex Machina}

### 2015-04-25 manual/projectrapport.tex
* added a small paragraph about the logger - - -> Do I need to provide more information there or is this enough?

\begin{figure}[h!]
	\makebox[\textwidth][c]{\includegraphics[width=\textwidth]{network.png}}
	\caption{Network cooperation diagram}
	\label{networkdesign}
\end{figure} 
Early benchmarks showed a repeatable throughput of 5e5 msg/s, with 4 kernels on modest hardware. The problems of (unintentional) out of order receipt and potential loss of messages are a non-issue in this context, since we use FIFO queues and loss of a message would implicate severe system failure (in which case message loss is the lesser problem).\\
Another advantage of hiding the message exchange behind a shared Network class instance is that, should this be required in the future, the kernel could run on different machines without requiring much redesigning of the system. An RPC implementation would simply have to implement the Network class on both sides.

@@

\chapter{Obstacles}
\section{Virtualization}
Virtualization proved a significant hurdle by exposing potential faults. Standard compliant code was found to be leaking memory by not calling container member destructors but only on a VM with Linux as Guest and Windows as Host. This was never reproducible in another setting.
\section{Threading}
libstdc++'s std::string implementation appears to be still Copy On Write (despite 21.4.1 p6 from the C++ standard) \cite{cpp}, triggering an alert by threadsanitizer for a possible race whenever a string was supposedly copied but internally still stored as a reference to the same object. Explicitly invoking the string constructor with the char[] proved to be a safe workaround here. Note that this problem is a known issue \cite{cow}.
\section{Toolchains}
To develop we used the required toolchains:\\ g++/clang++ as compiler, Eclipse+CDT as IDE, git for dvcs, jenkins as ci, cmake as build tool supplemented with some shell and python scripting.
\section{Logger}
The recommended logger (g3log), while powerful and fast, proved a dead end since it does not compile on Cygwin. 
Stijn therefore wrote his own implementation. %TODO Stijn
\chapter{Performance Analysis}
Whenever we made big design/architecture decisions, we first wrote some prototyping testcode to prove the viability (e.g. Network, Tracing, Logger).
\section{First comparison}
We use perf to get statistically sound benchmarks. The command run was :
\lstinline!$perf stat -r {rounds} {command}!\\

>>>>

\begin{figure}[h!]
	\makebox[\textwidth][c]{\includegraphics[width=\textwidth]{network.png}}
	\caption{Network cooperation diagram}
	\label{networkdesign}
\end{figure} 
Early benchmarks showed a repeatable throughput of 5e5 msg/s, with 4 kernels on modest hardware. The problems of (unintentional) out of order receipt and potential loss of messages are a non-issue in this context, since we use FIFO queues and loss of a message would implicate severe system failure (in which case message loss is the lesser problem).\\
Another advantage of hiding the message exchange behind a shared Network class instance is that, should this be required in the future, the kernel could run on different machines without requiring much redesigning of the system. An RPC implementation would simply have to implement the Network class on both sides.
\includegraphics[scale=0.2]{network.png}

@@

\chapter{Obstacles}
\section{Virtualization}
Virtualization proved a significant hurdle by exposing potential faults. Standard compliant code was found to be leaking memory by not calling container member destructors but only on a VM with Linux as Guest and Windows as Host. This was never reproducible in another setting.
\section{Threading}
libstdc++'s std::string implementation appears to be still Copy On Write (despite 21.4.1 p6 from the C++ standard) \cite{cpp}, triggering an alert by threadsanitizer for a possible race whenever a string was supposedly copied but internally still stored as a reference to the same object. Explicitly invoking the string constructor with the char[] proved to be a safe workaround here. Note that this problem is a known issue \cite{cow}.
\section{Toolchains}
To develop we used the required toolchains:\\ g++/clang++ as compiler, Eclipse+CDT as IDE, git for dvcs, jenkins as ci, cmake as build tool supplemented with some shell and python scripting.
\section{Logger}
The recommended logger (g3log), while powerful and fast, proved a dead end since it does not compile on Cygwin. 
Stijn therefore wrote his own implementation.\\
This implementation only covers the most basic needs:
\begin{itemize}
	\item asynchronous output
	\item thread safe
	\item multiple logging levels currently, the following 4 levels are supported:\\
		\emph{INFO}, \emph{DEBUG}, \emph{WARNING} and \emph{ERROR}
	\item levels can be individually switched on/off at compile time
	\item No logging code is executed whatsoever if the logger is disabled
	\item log messages contain filename and line number of the log command
\end{itemize}
It therefore lacks certain key features such as
\begin{itemize}
	\item crash safe
	\item custom assert that will log an error before shutting down
\end{itemize}
Chances are that we will wait with implementing these features until we actually need them.
\chapter{Performance Analysis}
Whenever we made big design/architecture decisions, we first wrote some prototyping testcode to prove the viability (e.g. Network, Tracing, Logger).
\section{First comparison}
We use perf to get statistically sound benchmarks. The command run was :
\lstinline!$perf stat -r {rounds} {command}!\\
